---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "man/figures/README-")
```

# Installation
**genet** can be installed via the `install_github` function in the **devtools** package.

```{r, eval = FALSE}
library(devtools)
install_github("aagrande/genet")
```

# Example: generating scale-free networks

Before invoking the generator via `genet`, we define:

* the number of communities `K` and node types `L`;
* a target community structure as encoded by the community assignments;
* a target degree distribution as encoded by a degree sequence with target degrees for each node type.

Depending on the application of interest, users can input their custom community structure. Here, for simplicity, we will generate a balanced structure via the `generate_balanced_assignments` function. This function randomly assigns each node type to `K_each` communities such that: no types have identical assignments; all communities have at least one type; and the resulting graph of type-to-type interactions is connected.

```{r}
library(genet)
K <- 100
L <- 200
Z <- generate_balanced_assignments(K, L, K_each = 5)
```

For the target degree distribution, we choose a scale-free distribution: $p(d)\propto d^{-\alpha}$, $d\in\{d_\text{min}, d_\text{min} + 1, \dots, d_\text{max}\}$. We choose the minimum and maximum degree, as well as the power law's exponent $\alpha$, so that the expected degree is approximately 15.

```{r}
d_min <- 3
d_max <- 500
alpha <- 1.937
as.numeric(dpower_law(3:500, alpha, d_min, d_max) %*% 3:500)
```

To closely reproduce the power-law distribution, we set the target degree sequence equal to the $(L + 1)$-quantiles of the distribution. Also, we adjust the type cardinalities so that the empirical distribution complies with the power law.

```{r, warning = FALSE}
target_degrees <- qpower_law(seq(0, 1, length.out = L), alpha, d_min, d_max)
probs <- dpower_law(target_degrees, alpha, d_min, d_max)
ratios <- probs / min(probs)
cards <- round(ratios * 2)
```

We can now generate networks with the desired community structure and target degree centrality distribution. We set `dc = TRUE` to specify that we target the degree centrality (as opposed to the eigencentrality).

```{r}
g <- genet(n = 10, assignments = Z, cards = cards, multiedges = TRUE, dc = TRUE,
           target_centrality = target_degrees)
```

The output of `genet` is a list of `igraph` graph objects. The first graph has the following number of nodes and edges:

```{r}
cat("# of nodes = ", igraph::vcount(g[[1]]), "\n# of edges = ", igraph::ecount(g[[1]]), sep = "")
```

Its empirical degree distribution is plotted below (the target power-law is represented in blue).

```{r empirical-distribution, echo = FALSE}
degree_distr <- table(igraph::degree(g[[1]]))
degree_distr <- degree_distr[degree_distr != 0]
degrees <- as.numeric(names(degree_distr))
counts <-  as.numeric(degree_distr)
max_d <- max(degrees)

compute_ticks <- function(x) {
  ticks <- (1:9) * 10 ^ x
  return(ticks[ticks <= max_d])
}

compute_labels <- function(ticks) {
  labs <- ticks
  labs[!labs %in% (10 ^ (0:5))] <- NA
  labs[1] <- 1
  labs[length(ticks)] <- ticks[length(ticks)]
  return(labs)
}

ticks <- unlist(sapply(0:5, compute_ticks))
labs <- compute_labels(ticks)
plot(log(degrees, base = 10), counts,
     ylab = "number of nodes", xlab = "degree", xaxt="n", pch = 16, cex = .7)
axis(side = 1, at = log(ticks, base = 10), labels = FALSE)
text(x = log(ticks, base = 10),  par("usr")[3], labels = labs, pos = 1, xpd = T)

lines(log(3:500, base = 10),
      dpower_law(3:500, alpha, d_min, d_max) * sum(degree_distr),
      col = "blue")
```

# References
- Grande and Feng, 2023. Principled and Scalable Network Generation with Built-in Community Structure and Target Centrality Measures.
