---
title: "Generating scale-free networks via `genet`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{genet-scale-free}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this vignette, we show how to generate graphs with an overlapping community structure and a target degree sequence drawn from a power-law distribution.

```{r setup}
library(genet)
```

First, we set set the number of communities `K` and types `L`, and fix a target community structure. Depending on the application of interest, users can input their custom community structure. Here, for simplicity, we will generate a balanced structure via the `generate_balanced_assignments` function. This function randomly assigns each type to `K_each` communities such that: no types have identical assignments; all communities have at least one type; and the resulting graph of type-to-type interactions is connected.

```{r}
K <- 100
L <- 200
Z <- generate_balanced_assignments(K, L, K_each = 5)
```

Next, we draw the target degree centrality sequence from a power-law degree sequence with the following parameters:

```{r}
d_min <- 3
d_max <- 500
alpha <- 1.937
```

We set the minimum and maximum degree to `r d_min` and `r d_max`, and the power law's exponent to `r alpha` so that the expected degree is approximately 15.

```{r}
as.numeric(dpower_law(3:500, alpha, d_min, d_max)  %*% 3:500)
```

To more closely reproduce the power-law distribution, we set the target degrees equal to the $(L + 1)$-quantiles of the distribution and then adjust the type cardinalities so that the empirical distribution complies with the power law.

```{r, warning = FALSE}
target_degrees <- qpower_law(seq(0, 1, length.out = L), alpha, d_min, d_max)
probs <- dpower_law(target_degrees, alpha, d_min, d_max)
ratios <- probs / min(probs)
cards <- round(ratios * 2)
```

We can now generate a random network.

```{r}
g <- genet(1, Z, cards, multiedges = T, dc = T, target_centrality = target_degrees)[[1]]
```

The output of `genet` is a list of `igraph` graph objects. 

```{r}
igraph::vcount(g)
igraph::ecount(g)
```

We can use the `igraph::degree` function to extract the empirical degree distribution of the graph. 

```{r}
degree_distr <- table(igraph::degree(g))
```

The resulting empirical distribution is plotted below. 

```{r, echo = FALSE}
degree_distr <- degree_distr[degree_distr != 0]
degrees <- as.numeric(names(degree_distr))
counts <-  as.numeric(degree_distr)
max_d <- max(degrees)

compute_ticks <- function(x) {
  ticks <- (1:9) * 10 ^ x
  return(ticks[ticks <= max_d])
}

compute_labels <- function(ticks) {
  labs <- ticks
  labs[!labs %in% (10 ^ (0:5))] <- NA
  labs[1] <- 1
  labs[length(ticks)] <- ticks[length(ticks)]
  return(labs)
}

ticks <- unlist(sapply(0:5, compute_ticks))
labs <- compute_labels(ticks)
plot(log(degrees, base = 10), counts,
     ylab = "number of nodes", xlab = "degree", xaxt="n", pch = 16, cex = .7)
axis(side = 1, at = log(ticks, base = 10), labels = FALSE)
text(x = log(ticks, base = 10),  par("usr")[3], labels = labs, pos = 1, xpd = T)
```
